Reduced growth due to belowground sink limitation is not fully explained by reduced photosynthesis

### 2016-07-15 manuscript/manuscript_pve.Rmd
very minor edits

\
COURTNEY E. CAMPANY^1^, BELINDA MEDLYN^1^ and REMKO A. DUURSMA^1^

\
^1^ Hawkesbury Institute for the Environment, University of Western Sydney, Locked Bag 1797, Penrith, NSW, Australia

\
Corresponding author: CE Campany, email: courtneycampany@gmail.com, telephone: +61 02 4570 1421, fax: +61 02 4570 1103

@@

- We manipulated soil volume to impose sink limitation of growth in *Eucalyptus tereticornis* seedlings. Seedlings were grown in containers of different sizes and planted flush to the soil alongside freely-rooted seedlings.

- Container volume negatively affected aboveground growth throughout the experiment, but mass partitioning to leaves, stems and roots was unchanged after 120 days. Photosynthetic capacity was significantly reduced in containerized seedlings, and was related to both reduced leaf nitrogen content and starch accumulation. However, the reduction in net leaf photosynthesis over the growth period was insufficient to explain the reduction in growth, so that we also observed a reduction in whole-plant C use efficiency (CUE) with container size.

@@

\
In the absence of a good understanding of the effects of sink limitation on physiological processes, most attempts to model sink limitation of growth to date have done so by abandoning the principle of mass balance [e.g. @fatichi2013reconciling; @fatichi2014moving]. This approach is unsatisfactory because mechanistic models should obey the laws of physics. Growth models must take into account several simultaneous processes affecting the C balance at any given time [@fourcaud2008plant], thus incomplete modelling approaches often lead to unaccounted for C. Our goal in this paper was to investigate how the C balance of plants is modified by sink limitation, by experimentally imposing a sink limitation to growth.

\
In previous studies, the sink limitation of growth has been explored by manipulations of both C source and C sink activity. Examples of experimental approaches that manipulate C source activity include elevated atmospheric CO2 concentration and partial defoliation experiments. Elevated CO~2~ causes an increase in leaf-level photosynthetic rate, *A*~n~ [@drake1997more; @ainsworth2007response]. This increased photosynthetic uptake is often observed to drive an increase in plant biomass production [@norby2005forest]. Evidence from a wide range of elevated CO~2~ experiments, however, also reveals that even with an average photosynthetic enhancement of over 30 %, the biomass growth rate only increases by around 10 % [@kirschbaum2011does]. This implies that other process of tree C balance must be affected, which has been noted in CO~2~ manipulation studies that attempt to close the C budget [@wang1998carbon; @hamilton2002forest; @schafer2003exposure].

\
In partial defoliation experiments, source activity is decreased by reduction of photosynthesizing leaf area. In these experiments, compensatory increases in *A*~n~ of the remaining foliage are commonly found, and have been attributed to several processes including indirect effects on resource availability, such as increased stomatal conductance [@layne1995end] or enhanced leaf nutrient status [@turnbull2007increased; @pinkard2011gas], or more direct effects of modified sink-source balance such as, reduction in end-product inhibition [@iglesias2002regulation; @zhou2003changes; @handa2005test], enhanced biochemical activity [@ovaska1993b;@layne1995end;@pinkard2011gas], and regulatory sugar signaling [@eyles2013whole]. Similar to elevated CO~2~ experiments, these increases in *A*~n~ in defoliation experiments do not increase tree growth [@palacio2012fast; @markkola2004defoliation; @ovaska1993a]. As these studies do not attempt to scale the degree of compensatory response in leaf C gain to the entire canopy, the overall affect of source manipulation on tree C balance and growth is still unclear.

@@

\
Our goal in this paper was to impose a sink limitation of growth, by manipulating container size and thus rooting volume, and to examine how the C balance of the plants was affected. We grew *Eucalyptus tereticornis* seedlings in a range of container sizes, using freely-rooted seedlings as a control for the container size treatments. Seedlings were maintained under well-watered conditions in order to isolate the effect of restricted soil volume from that of low water availability. We tracked leaf-level gas exchange, leaf carbohydrate accumulation and seedling allometry over the course of 4 months. We drove a simple whole-plant photosynthesis model with measured leaf gas exchange and interpolated leaf area data to estimated net canopy C uptake over the experimental period, and compared the result with total seedling biomass. We anticipated that both photosynthesis and growth would be reduced under belowground sink limitation. The principal questions that we addressed were: Firstly, which physiological process best explains the down-regulation of photosynthesis? Secondly, does this down-regulation fully explain the observed response of growth, or are other components of the C balance also affected.

@@

\
Forty-nine *Eucalyptus tereticornis* seedlings, 20 weeks old and approximately 40 cm tall in tube stock, were chosen from a single local Cumberland plain cohort. Six additional seedlings were harvested before planting to measure initial leaf area and dry mass of leaves, stems and roots. Previous container experiments have confirmed that species with tap roots (similar to *E. tereticornis*) use the center of the container as the medium for thick roots, leaving the periphery of the soil as the most active sites for fine root proliferation [@biran1980a; @biran1980b]. By using a species with taproot growth and manipulations of container length rather than width, we believed that a more realistic test of growth inhibition through constrained soil volume would be achieved.

@@

\
Plants were watered weekly or when needed to maintain soil moisture at field capacity (13-15 %). Drain systems were built into each pot to prevent pooling of water throughout the experiment. Pooling of water could lead to an anaerobic environment around the root that could hinder the uptake of water through reduced root conductance [@poorter2009causes], an undesired experimental artifact. A collection compartment in the bottom of containers, containing gravel covered by root exclusion mesh, was used to collect excess water for 20, 25, and 35 l containers. Plastic tubing (6 mm diameter) was inset into the gravel layer and extended through the top of the container. A lysimeter pump was then used to suction excess water, through the tubing, as needed. For small containers (5, 10, and 15 L) a simple bottom plug was used to drain excess water from the gravel compartment. Each container was inspected after every watering and rainfall event to determine if pooling had occurred.

@@

##Leaf water potential
Predawn ($\Psi$~pd~) and midday ($\Psi$~l~) leaf water potentials were measured for each seedling using a PMS 1505D pressure chamber (PMS Instruments, Albany, OR, USA) on fully expanded leaves during the same time period as AC~i~ and *R*. Leaves were detached and immediately stored inside foil-covered bags before water potential measurements were performed. $\Psi$~pd~ was measured before sunrise and $\Psi$~l~ at 13:00 h (&pm30 m). These measurements were used as a measure of static water stress on the seedlings [@sellin1999does] and to ensure that the bulk soil water availability was high enough for plants to avoid water stress as they became larger and roots filled the soil volume.

@@

\
Both *J*~max~ and *V*~cmax~ were significantly higher in free seedlings (30 % and 26 %, respectively) than container-grown seedlings with little variation between container volume treatments (Table `r pvetab3`). Overall, the J~max~:V~cmax~ ratio was relatively conserved across all treatments (1.6&pm;0.02), which is consistent with findings across many tree species [@medlyn1996optimal; @leuning1997scaling; @medlyn1999effects; @warren2003photosynthesis; @crous2008elevated]. Leaf dark respiration rates were not significantly different across soil volumes (Table `r pvetab3`). The *g*~1~ parameter, generated for each seedling from the Medlyn *et al.* [-@medlyn2011reconciling] optimal stomatal conductance model, was lowest in the free seedling treatment and was marginally different across soil volume treatments (Table `r pvetab3`).

\
Neither $\Psi$~pd~ nor $\Psi$~l~ were different across treatments, with mean values of -0.27 and -1.2 MPa across all seedlings, respectively. Although *g*~s~ in free seedlings was generally higher than those in containers (Table `r pvetab3`, *P* < 0.001), the mean rates for all seedlings were high at 0.37 mol m^-2^ s^-1^ and did not change throughout the course of the experiment. Additionally, leaf $\delta$^13^C at final harvest was not different across treatments (Table `r pvetab1`). Combined, these indices provide strong evidence that water stress was not apparent on these seedlings throughout the experiment. Soil N at harvest was not different across soil volumes ($\bar{x}$ = .045 %), with minimal decreases from pre-planting values ($\bar{x}$ = .049 %). This similarity indicates that nutrient leaching from free seedlings or from draining of containers following natural rainfall events did not differ between treatments.

@@

#Discussion
This study utilized a novel field design to manipulate belowground sink limitation and physically restrict *Eucalyptus tereticornis* seedling biomass production. Our goal was to identify how sink limitation of growth modified the C balance of seedlings. We found a strong reduction of growth in containerized plants. Leaf *A*~n~ was also strongly reduced, but this reduction was insufficient to fully explain the reduction in growth, implying that components such as whole-plant respiratory loss were also likely affected.

@@

##Biomass partitioning under sink limitation
As biomass partitioning is likely controlled by the source and sink strength of all organs [@poorter2012biomass], it was important to determine which tissue components were most affected by the container size treatments. It was necessary to distinguish if growth was affected beyond ontogenetic constraints, by correcting for size, as biomass distribution is strongly size-dependent [@gould1966allometry; @lleonart2000removing]. In this study, there was no significant difference in root, leaf, or stem biomass partitioning with reduced soil volume compared to free seedlings, once ontogenetic drift was taken into account (Figure `r allocfig`a,b). Although shifts in partitioning have been previously reported, specifically for nutrient limitation [@mcconnaughay1999biomass, and references therein], a constant ratio of fine root mass to leaf mass was observed in this study across all treatments. These results suggest that seedlings kept a conservative strategy to maintain homeostasis with with biomass partitioning, instead of a functional partitioning response to optimize limiting soil resources.

\
The lack of detected shifts in partitioning to fine roots provides evidence against an optimal foraging strategy for seedlings in containers. It is instead possible that lateral root development is affected by inanimate obstacles and avoiding growth towards container walls may improve resource gain [@falik2005root]. Root exudation may have increased with reduced rooting volumes to facilitate N uptake in favor of increasing partitioning to root biomass. The sensitivity of roots to their own exudates near obstructions may also prevent further growth [@semchenko2008foraging]. Here, we found that FRLD was highest in smallest containers suggesting that root restriction likely occurred as simple function of available rooting space. Additionally, physical restriction of root proliferation could have impacted root development and morphology prior to shifts in mass partitioning. Here, increases in SRL were detected in several of the soil volume treatments. This is not surprising as plants in containers have been shown to have different root morphology to field grown plants [@nesmith1998effect]. The poor soil quality used in our experiment and root restriction, however, likely decreased the capacity of this morphological response to increase N uptake.

##Do reductions in photosynthesis explain reductions in seedling growth?
We use a simple approach to estimate cumulative seedling C gain with measured reductions in leaf *A*~n~ , via soil volume effects.  Although reductions in both *A*~max~ and biomass were strongly correlated among treatments, as hypothesized by Poorter *et al.* [-@poorter2012pot], we provide evidence that the negative effects of sink limitation on *A*~n~ do not fully explain reduced seedling growth. As whole-plant C use efficiency was decreased in seedlings with belowground sink limitation, this suggests that other components of the C balance were affected. These results are especially noteworthy for process-based growth models that parameterize CUE, tissue respiration and C allocation as fixed processes. For example, classical approaches in tree growth and production modelling are often driven by inputs of C assimilation and processes such as respiration are considered proportional to biomass [@le2001carbon] or growth rate [@tjoelker1999acclimation]. It is possible these results represent an initial overestimation of *A*~n~, however, the robust empirical based methods used to generate photosynthetic parameters (*J*~max~, *Vc*~max~, *R*~dark~ and *g*~1~) make this unlikely. Instead, our results indicate a need to evaluate how oversimplified representations of processes other than *A*~n~ affect models which distinguish the fate of assimilate C within a plant. Doing so will provide valuable input to future models as assimilate allocation is a key component in C-balance driven plant growth models [@lacointe2000carbon].

\
Focusing only on empirical results in this study points to *A*~n~ driving reductions in seedling growth. Modelling cumulative net seedling C gain, however, reveals that other components of the C balance, beyond *A*~n~, are required to explain the observed seedling biomass response to sink manipulation. This finding is noteworthy, as partial accounting of the different mechanisms involved in plant C balance can lead to erroneous conclusions [@valentini2000respiration]. For example, the fraction of photosynthate used in respiration is known to vary depending on species and local environmental conditions [@Lambers2008], yet is often considered a static parameter in process based growth models. Thus, we agree with conclusions from Valentine and M&#228;kel&#228; [-@valentine2005bridging] where the problem with predicting tree growth is a problem in forecasting the assimilation and allocation of C and other constituents. Future empirical and modelling studies should focus on how feedbacks from sink activity affect both rates of *A*~n~ and the fate of C allocated to growth, respiration and C storage in leaves. However, the degree to which these mechanisms regulate the C available for growth will undoubtedly shift across different species and environmental conditions.

>>>>

\
Courtney E. Campany^1^, Belinda Medlyn^1^ and Remko A. Duursma^1^

\
^1^ Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith, NSW, Australia

\
Corresponding author: C.E. Campany, email: courtneycampany@gmail.com, telephone: +61 02 4570 1421, fax: +61 02 4570 1103

@@

- We manipulated soil volume to impose sink limitation of growth in *Eucalyptus tereticornis* seedlings. Seedlings were grown in the field in containers of different sizes and planted flush to the soil alongside freely-rooted seedlings.

- Container volume negatively affected aboveground growth throughout the experiment, but after 120 days, biomass distribution in  leaves, stems and roots was not different. Photosynthetic capacity was significantly reduced in containerized seedlings, and was related to both reduced leaf nitrogen content and starch accumulation. However, the reduction in net leaf photosynthesis over the growth period was insufficient to explain the reduction in growth, so that we also observed a reduction in whole-plant C use efficiency (CUE) with container size.

@@

\
In the absence of a good understanding of the effects of sink limitation on physiological processes, most attempts to model sink limitation of growth to date have done so by abandoning the principle of mass balance [e.g. @fatichi2013reconciling; @fatichi2014moving]. This approach is unsatisfactory because mechanistic models should obey the laws of physics. Growth models must take into account several simultaneous processes affecting the C balance at any given time [@fourcaud2008plant], thus incomplete modelling approaches often lead to unaccounted-for C. Our goal in this paper was to investigate how the C balance of plants is modified by sink limitation, by experimentally imposing a sink limitation to growth.

\
In previous studies, the sink limitation of growth has been explored by manipulations of both C source and C sink activity. Examples of experimental approaches that manipulate C source activity include elevated atmospheric CO~2~ concentration and partial defoliation experiments. Elevated CO~2~ causes an increase in leaf-level photosynthetic rate, *A*~n~ [@drake1997more; @ainsworth2007response]. This increased photosynthetic uptake is often observed to drive an increase in plant biomass production [@norby2005forest]. Evidence from a wide range of elevated CO~2~ experiments, however, also reveals that even with an average photosynthetic enhancement of over 30 %, the biomass growth rate only increases by around 10 % [@kirschbaum2011does]. This implies that other process of tree C balance must be affected, which has been noted in CO~2~ manipulation studies that attempt to close the C budget [@wang1998carbon; @hamilton2002forest; @schafer2003exposure].

\
In partial defoliation experiments, source activity is decreased by a reduction of photosynthesizing leaf area. In these experiments, compensatory increases in *A*~n~ of the remaining foliage are commonly found, and have been attributed to several processes including indirect effects on resource availability, such as increased stomatal conductance [@layne1995end] or enhanced leaf nutrient status [@turnbull2007increased; @pinkard2011gas], or more direct effects of modified sink-source balance such as, reduction in end-product inhibition [@iglesias2002regulation; @zhou2003changes; @handa2005test], enhanced biochemical activity [@ovaska1993b;@layne1995end;@pinkard2011gas], and regulatory sugar signaling [@eyles2013whole]. Similar to elevated CO~2~ experiments, these increases in *A*~n~ in defoliation experiments do not increase tree growth [@palacio2012fast; @markkola2004defoliation; @ovaska1993a]. As these studies do not attempt to scale the degree of compensatory response in leaf C gain to the entire canopy, the overall affect of source manipulation on tree C balance and growth is still unclear.

@@

\
Our goal in this paper was to impose a sink limitation of growth, by manipulating container size and thus rooting volume, and to examine how the C balance of the plants was affected. We grew *Eucalyptus tereticornis* seedlings in a range of container sizes in field conditions, using freely-rooted seedlings as a control for the container size treatments. Seedlings were maintained under well-watered conditions in order to isolate the effect of restricted soil volume from that of low water availability. We tracked leaf-level gas exchange, leaf carbohydrate accumulation and seedling allometry over the course of 4 months. We drove a simple whole-plant photosynthesis model with measured leaf gas exchange and interpolated leaf area data to estimated net canopy C uptake over the experimental period, and compared the result with total seedling biomass. We anticipated that both photosynthesis and growth would be reduced under belowground sink limitation. The principal questions that we addressed were: 1) which physiological process best explains the down-regulation of photosynthesis?, and 2) does this down-regulation fully explain the observed response of growth, or are other components of the C balance also affected?

@@

\
Forty-nine *Eucalyptus tereticornis* seedlings, 20 weeks old and approximately 40 cm tall in tube stock, were chosen from a single local Cumberland plain (western Sydney) cohort. Six additional seedlings were harvested before planting to measure initial leaf area and dry mass of leaves, stems and roots. Previous container experiments have confirmed that species with tap roots (similar to *E. tereticornis*) use the center of the container as the medium for thick roots, leaving the periphery of the soil as the most active sites for fine root proliferation [@biran1980a; @biran1980b]. By using a species with taproot growth and manipulations of container length rather than width, we believed that a more realistic test of growth inhibition through constrained soil volume would be achieved.

@@

\
Plants were watered weekly or when needed to maintain soil moisture at field capacity (13-15 %). Drain systems were built into each pot to prevent pooling of water throughout the experiment. Pooling of water could lead to an anaerobic environment around the root that could hinder the uptake of water through reduced root conductance [@poorter2009causes], an undesired experimental artefact. A collection compartment in the bottom of containers, containing gravel covered by root exclusion mesh, was used to collect excess water for 20, 25, and 35 l containers. Plastic tubing (6 mm diameter) was inset into the gravel layer and extended through the top of the container. A lysimeter pump was then used to suction excess water, through the tubing, as needed. For small containers (5, 10, and 15 L) a simple bottom plug was used to drain excess water from the gravel compartment. Each container was inspected after every watering and rainfall event to determine if pooling had occurred.

@@

##Leaf water potential
Predawn ($\Psi$~pd~) and midday ($\Psi$~l~) leaf water potentials were measured for each seedling using a PMS 1505D pressure chamber (PMS Instruments, Albany, OR, USA) on fully expanded leaves during the same time period as AC~i~ and *R*. Leaves were detached and immediately stored inside foil-covered bags before water potential measurements were performed. $\Psi$~pd~ was measured before sunrise and $\Psi$~l~ at 13:00 h (&pm; 30 min). These measurements were used as a measure of static water stress on the seedlings [@sellin1999does] and to ensure that the bulk soil water availability was high enough for plants to avoid water stress as they became larger and roots filled the soil volume.

@@

\
Both *J*~max~ and *V*~cmax~ were significantly higher in free seedlings (30 % and 26 %, respectively) than container-grown seedlings with little variation between container volume treatments (Table `r pvetab3`). Overall, the *J*~max~ / *V*~cmax~ ratio was conserved across all treatments (1.6&pm;0.02), which is consistent with findings across many tree species [@medlyn1996optimal; @leuning1997scaling; @medlyn1999effects; @warren2003photosynthesis; @crous2008elevated]. Leaf dark respiration rates were not significantly different across soil volumes (Table `r pvetab3`). The *g*~1~ parameter, generated for each seedling from the Medlyn *et al.* [-@medlyn2011reconciling] optimal stomatal conductance model, was lowest in the free seedling treatment and was marginally different across soil volume treatments (Table `r pvetab3`).

\
Neither $\Psi$~pd~ nor $\Psi$~l~ were different across treatments, with mean values of -0.27 and -1.2 MPa across all seedlings, respectively. Although *g*~s~ in free seedlings was generally higher than those in containers (Table `r pvetab3`, *P* < 0.001), the mean rates for all seedlings were high at 0.37 mol m^-2^ s^-1^ and did not change throughout the course of the experiment. Additionally, leaf $\delta$^13^C at final harvest was not different across treatments (Table `r pvetab1`). Combined, these indices provide strong evidence that water stress was not apparent on these seedlings throughout the experiment. Soil N at harvest was not different across soil volumes (mean of .045 %), with minimal decreases from pre-planting values (mean of = .049 %). This similarity indicates that nutrient leaching from free seedlings or from draining of containers following natural rainfall events did not differ between treatments.

@@

#Discussion
This study utilized a novel field design to manipulate belowground sink limitation and physically restrict *Eucalyptus tereticornis* seedling growth. Our goal was to identify how sink limitation of growth modified the C balance of seedlings. We found a strong reduction of growth in containerized plants. Leaf *A*~n~ was also reduced, but this reduction was insufficient to fully explain the reduction in growth, implying that components such as whole-plant respiratory loss were also likely affected.

@@

##Biomass partitioning under sink limitation
As biomass partitioning is likely controlled by the source and sink strength of all organs [@poorter2012biomass], it was important to determine which tissue components were most affected by the container size treatments. It was necessary to distinguish if growth was affected beyond ontogenetic constraints, by correcting for size, as biomass distribution is strongly size-dependent [@gould1966allometry; @lleonart2000removing]. In this study, there was no significant difference in root, leaf, or stem biomass partitioning with reduced soil volume compared to free seedlings, once ontogenetic drift was taken into account (Figure `r allocfig`a,b). Although shifts in partitioning have been previously reported, specifically for nutrient limitation [@mcconnaughay1999biomass, and references therein], a constant ratio of fine root mass to leaf mass was observed in this study across all treatments. These results suggest that seedlings kept a conservative strategy to maintain homeostasis with biomass partitioning, instead of a functional partitioning response to optimize limiting soil resources.

\
The lack of detected shifts in partitioning to fine roots provides evidence against an optimal foraging strategy for seedlings in containers. It is instead possible that lateral root development is affected by inanimate obstacles and avoiding growth towards container walls may improve resource gain [@falik2005root]. Root exudation may have increased with reduced rooting volumes to facilitate N uptake in favor of increasing partitioning to root biomass. The sensitivity of roots to their own exudates near obstructions may also prevent further growth [@semchenko2008foraging]. Here, we found that fine root length density was highest in smallest containers suggesting that root restriction likely occurred as simple function of available rooting space. Additionally, physical restriction of root proliferation could have impacted root development and morphology prior to shifts in mass partitioning. Here, increases in specific root length were detected in several of the soil volume treatments. This is not surprising as plants in containers have been shown to have different root morphology to field grown plants [@nesmith1998effect]. The poor soil quality used in our experiment and root restriction, however, likely decreased the capacity of this morphological response to increase N uptake.

##Do reductions in photosynthesis explain reductions in seedling growth?
We use a simple approach to estimate cumulative seedling C gain with measured reductions in leaf *A*~n~ , via soil volume effects.  Although reductions in both *A*~max~ and biomass were strongly correlated among treatments, as hypothesized by Poorter *et al.* [-@poorter2012pot], we provide evidence that the negative effects of sink limitation on *A*~n~ do not fully explain reduced seedling growth. As whole-plant C use efficiency was decreased in seedlings with belowground sink limitation, this suggests that other components of the C balance were affected. These results are especially noteworthy for process-based growth models that parameterize CUE, tissue respiration and C allocation as fixed processes. For example, classical approaches in tree growth and production modelling are often driven by inputs of C assimilation and processes such as respiration are considered proportional to biomass [@le2001carbon] or growth rate [@tjoelker1999acclimation]. It is possible these results represent an initial overestimation of *A*~n~, however, the robust empirical based methods used to generate photosynthetic parameters (*J*~max~, *V*~cmax~, *R*~dark~ and *g*~1~) make this unlikely. Instead, our results indicate a need to evaluate how oversimplified representations of processes other than *A*~n~ affect models which distinguish the fate of assimilate C within a plant. Doing so will provide valuable input to future models as assimilate allocation is a key component in C-balance driven plant growth models [@lacointe2000carbon].

\
Focusing only on empirical results in this study points to *A*~n~ driving reductions in seedling growth. Modelling cumulative net seedling C gain, however, reveals that other components of the C balance, beyond *A*~n~, are required to explain the observed seedling biomass response to sink manipulation. This finding is noteworthy, as partial accounting of the different mechanisms involved in plant C balance can lead to erroneous conclusions [@valentini2000respiration]. For example, the fraction of photosynthate used in respiration is known to vary depending on species and local environmental conditions [@Lambers2008], yet is often considered a static parameter in process based growth models. Thus, we agree with conclusions from Valentine and M&#228;kel&#228; [-@valentine2005bridging] that the problem with predicting tree growth is a problem in forecasting the assimilation and allocation of C and other constituents. Future empirical and modelling studies should focus on how feedbacks from sink activity affect both rates of *A*~n~ and the fate of C allocated to growth, respiration and C storage in leaves. However, the degree to which these mechanisms regulate the C available for growth will undoubtedly shift across different species and environmental conditions.
